Effectiveness of community-based TURF-reserves in Mexican small-scale fisheries

### 2019-02-11 docs/manuscript.tex
Added changes for Rev1

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Small-scale fisheries face
great challenges since they are difficult to monitor, enforce, and
manage. Combining territorial use rights for fisheries (TURF) with
no-take marine reserves to create TURF-reserves can improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 12
years, 18 old and new community-based Mexican TURF-reserves gained legal
recognition thanks to a regulation passed in 2014; their effectiveness
has not been formally evaluated. We combine causal inference techniques
and the Social-Ecological Systems framework to provide a holistic
evaluation of community-based TURF-reserves in three coastal communities
in Mexico. We find that while reserves have not yet achieved their
stated goal of increasing the density of lobster and other benthic
invertebrates, they continue to receive support from the fishing
communities. A lack of clear ecological and socioeconomic effects likely
results from a combination of factors. First, some of these reserves
might be too young for the effects to show. Second, the reserves are not
large enough to protect mobile species, like lobster. Third, variable
and extreme oceanographic conditions have impacted harvested
populations. Fourth, local fisheries are already well managed, and it is
unlikely that reserves might have a detectable effect in catches.
However, these reserves may provide a foundation for establishing
additional, larger marine reserves needed to effectively conserve mobile
species.

@@

Conservation science has shown how marine reserves lead to increased
biomass, species richness, and abundance within the protected regions
\citep{lester_2009}, and that these may have a series of additional
benefits such as climate change mitigation, protection from
environmental variability, and fisheries benefits
\citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise,
research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access and even similar
to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The
benefits resulting from reserves established within TURFs (\emph{i.e.}
TURF-reserves) should be captured exclusively by the group of fishers
with exclusive access \citep{gelcich_2015}. Although in theory these
systems are successful \citep{smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success. Moreover, TURF-reserve systems are inherently intricate
social-ecological systems, and their effectiveness must depend on how
environmental and social factors combine and interact
\citep{ostrom_2009,gelcich_2015}. It is therefore important to consider
not only the indicators of interest, but also the governance settings
under which the reserves operate.

Recent norms in fisheries regulation in Mexico provide a ripe
opportunity to study the effectiveness of community-based TURF-reserves
in small-scale fisheries. In Mexico, a norm created in 2014 allows
fishers to request legal recognition of community-based reserves as
``Fish Refuges'' (\emph{Zona de Refugio Pesquero}; \citet{nom}). Since
2012, old and new marine reserves have gained legal recognition as Fish
Refuges. Of these, 18 were originally implemented within TURFs. However,
their effectiveness has not yet been formally evaluated and reported in
the scientific literature.

@@

Before discussing our data collection methods and describing our
analyses, our case studies warrant some background. Community-based
marine reserves that are implemented within TURFs are a form of
TURF-reserve, voluntarily established and enforced by local communities.
This bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{beger_2004,smallhornwest_2018}. Community-based spatial closures
occur in different contexts, like the \emph{kapu} or \emph{ra'ui} areas
in the Pacific Islands \citep{johannes_2002,bohnsack_2004}. However,
community-based reserves can be hard to enforce if they are not legally
recognized. In such conditions, TURF fishers must rely on the exclusive
access of the TURF to maintain high levels of compliance.

@@

The most common setup of community-based TURF-reserves in Mexico is the
following. Fishers from a given community are assembled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and
each community usually hosts no more than one cooperative. The profits
from each TURF are shared amongst all fishers from the cooperative.
Fishing cooperatives interested in implementing marine reserves work
with CSOs to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves as Fish Refuges following a series of
studies outlined in the regulation \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserve systems
implemented in Mexican TURF-managed fisheries, therefore making them
TURF-reserves (Fig \ref{fig:map}A). The first one was created by the
\emph{Buzos y Pescadores de la Baja California} fishing cooperative,
located in Isla Natividad in the Baja California Peninsula (Fig
\ref{fig:map}B). The main fishery in the island is the spiny lobster
(\emph{Panulirus interruptus}), but other resources like finfish, sea
cucumber, sea urchin, snail, and abalone are also an important source of
income. In 2006, the community decided to implement two marine reserves
within their fishing grounds. The objective of these reserves was ``to
protect commercially important invertebrate species''; mainly lobster
and abalone. The reserves obtained legal recognition in 2018
\citep{dof_website_2018}.

@@

The biological data are collected by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record species richness
and abundances of fish and invertebrate species along replicate
transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and
control sites \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size
structures are also collected during fish surveys. All sites were
surveyed annually, and at least once before implementation of the
reserves. A summary of sampling effort is shown in the supplementary
materials (Tables S1-S2).

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for TURF-managed cooperatives with and without
marine reserves. In this case our treated unit are the cooperatives
(\emph{i.e.} communities) that have implemented a reserve within their
TURF, and the controls are nearby communities that have a TURF but did
not implement a reserve. Cooperatives incorporated in this analysis
belong to larger regional-level Cooperative Federations, and are exposed
to the same markets and institutional frameworks, making them plausible
controls \citep{mccay_2017,ayer_2018}. Landings and revenues were
aggregated at the cooperative-year level, and revenues were adjusted to
represent 2014 values by the Consumer Price Index for Mexico
\citep{oecd_2017}. A table with summary statistics for this data is
provided in the supplementary materials (Table S3, Figure S5).

@@

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio},
but in this case the \(Treated\) dummy variable indicates if the
community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)).
These regression models allow us to establish a causal link between the
implementation of marine reserves and the observed trends by accounting
for temporal and site-specific dynamics \citep{depalma_2018}. We fit one
model per community and indicators (\emph{e.g.} there are three models
for lobster density, one for each community) for a total of 12
biological model fits and four socioeconomic model fits. Model
coefficients were estimated via ordinary least-squares and used
heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All
analyses were performed in R version 3.5.2 and R Studio version 1.1.456
\citep{R_2018}. All data and code needed to reproduce our analyses are
available in a GitHub repository at:
\url{https://github.com/jcvdav/ReserveEffect}.

We use the SES framework to evaluate each community and create a
narrative that provides context for each community. The use of this
framework standardizes our analysis and allows us to communicate our
results in a common language across fields by using a set of previously
defined variables and indicators. We based our variable selection
primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who
operationalized and analyzed Mexican fishing cooperatives using this
framework, and identified the key variables relevant to fishing
cooperatives in Mexico. We also incorporate other relevant variables
known to influence reserve performance following
\citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table
\ref{table:ses} shows the selected variables, along with definitions and
values.

@@

Lobster landings and revenue were only available for Isla Natividad and
Maria Elena (Fig \ref{fig:lobsters}). For all years before
implementation, the effect sizes are close to zero, indicating that the
control and treatment sites have similar pre-treatment trends,
suggesting that these are plausible controls. However, effect sizes do
not change after the implementation of the reserve. Interestingly, the
negative effect observed for Isla Natividad on year 5 corresponds to the
2011 hypoxia events \citep{micheli_2012-EU}. The only positive change
observed in lobster landings is for Isla Natividad in 2014
(\(p < 0.1\)). The three years of post-implementation data for Maria
Elena do not show a significant effect of the reserve. Isla Natividad
shows higher revenues after the implementation of the reserve, as
compared to the control communities. However, these changes are only
significant for the third year (\(p < 0.05\)). Full tables with model
coefficients are presented in the supplementary materials (Tables
S4-S5).

@@

Our analysis of the SES (Table \ref{table:ses}) shows that all analyzed
communities share similarities known to foster sustainable resource
management and increase reserve effectiveness. For example, fishers
operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide
exclusive access to resources and reserves. Along with their relatively
small groups (A1 - Number of relevant actors), Isolation (A3),
Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated
sanctions (GS10.1), these fisheries have solid governance structures
that enable them to monitor their resources and enforce rules to ensure
sustainable management. In general, success of conservation initiatives
depends on the incentives of local communities to maintain a healthy
status of the resources upon which they depend \citep{jupiter_2017}. Due
to the clarity of access rights and isolation, the benefits of
conservation directly benefit the members of the fishing cooperatives,
which have favored the development of efficient community-based
enforcement systems. However, our SES analysis also highlights factors
that might hinder reserve performance or mask outcomes. While total
reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual
reserves are often small (RS3); the largest reserve is only 4.37 km
\textsuperscript{2}, and the smallest one is 0.09 km
\textsuperscript{2}. Reserves are also relatively young (RS5).
Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely
that marine reserves will result in increased catches.

@@

Our analyses of socioeconomic indicators has two limitations. First, we
only look at landings and revenues by landings for communities with and
without TURF-reserves. There are a number of other possible indicators
that could show a change due to the implementation of the reserve.
Notably, one often cited in the literature is additional benefits, such
as tourism \citep{viana_2017}. However, it is unlikely that the
evaluated communities will experience tourism benefits due to their
remoteness and the lack of proper infrastructure to sustain tourism. A
second limitation of our socioeconomic analysis is that we do not
observe effort data, which may mask the effect of the reserve. For
example, if catches remain relatively unchanged but fishing effort
decreased, that would imply a larger catch per unit effort and thus
higher profitability, provided that cost per unit effort does not
increase.

A first possible explanation for the lack of effectiveness may be the
young age of the reserves. Literature shows that age and enforcement are
important factors that influence reserve effectiveness
\citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest
reserves, and our SES analysis suggests that all communities have a
well-established community-based enforcement system. With these
characteristics, one would expect the reserves to be effective. Maria
Elena and Punta Herrero are relatively young reserves (\emph{i.e.}
\textless{} 6 years old) and effects may not yet be evident due to the
short duration of protection, relative to the life histories of the
protected species; community-based marine reserves in tropical
ecosystems may take six years or more to show a spillover effect
\citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}). Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}, however, abalone are less mobile than
lobsters, and perhaps the reserves provide enough protection to these
sedentary invertebrates, but not lobsters. Design principles developed
by \citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, fishers
may favor implementation of reserves that pose low fishing costs due to
their small size or location. Our analysis of economic data supports
this hypothesis, as neither landings nor revenues showed the expected
short-term reductions associated to the first years of reserve
implementation \citep{ovando_2016-Wg}.

@@

\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero). Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. The presented narrative applies equally for all communities unless otherwise noted.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12cm}}
\hline
Variable & Narrative\\
\hline
\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
\hline
\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates. Reserve placement is decided by the community. Fishers use GPS units and landmarks.\\
\hline
\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
\hline
\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
\hline
\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
\hline
\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
\hline
\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
\hline
\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each). Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
\hline
\multicolumn{2}{l}{\textbf{Actors (A)}}\\
\hline
\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
\hline
\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers. IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
\hline
\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
\hline
\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster. These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
\hline
\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates. These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
\hline
\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules. They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
\hline
\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds. Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
\hline
GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves. These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
\hline
\end{tabular}}
\end{table}

>>>>

Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. Small-scale fisheries face great challenges since they are difficult to monitor, enforce, and manage. Combining territorial use rights for fisheries (TURF) with no-take marine reserves to create TURF-reserves can improve the performance of small-scale fisheries by buffering fisheries from environmental variability and management errors, while ensuring that fishers reap the benefits of conservation investments. In the last 13 years, 18 old and new community-based Mexican TURF-reserves gained legal recognition thanks to a regulation passed in 2012; their effectiveness has not been formally evaluated. We combine causal inference techniques and the Social-Ecological Systems framework to provide a holistic evaluation of community-based TURF-reserves in three coastal communities in Mexico. We find that reserves have not yet achieved their stated goal of increasing the density of lobster and other benthic invertebrates. A lack of clear ecological and socioeconomic effects likely results from a combination of factors. First, some of these reserves might be too young for the effects to show. Second, the reserves are not large enough to protect mobile species, like lobster. Third, variable and extreme oceanographic conditions have impacted harvested populations. Fourth, local fisheries are already well managed, and it is unlikely that reserves might have a detectable effect in catches. However, these reserves may provide a foundation for establishing additional, larger marine reserves needed to effectively conserve mobile species.

@@

Conservation science has shown how marine reserves may lead to increased biomass, species richness, and abundance within the protected regions \citep{lester_2009}, and that these may have a series of additional benefits such as mitigation and adaptation to climate change effects, protection from environmental variability, and fisheries benefits \citep{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. Likewise, research on TURFs has shown that these areas have higher abundance of targeted species than sites operating under open access and even similar to that of marine reserves \citep{gelcich_2008,gelcich_2012}. The benefits resulting from reserves established within TURFs (\emph{i.e.} TURF-reserves) should be captured exclusively by the group of fishers with exclusive access \citep{gelcich_2015}. Although in theory these systems are expected to be successful \citep{smallhornwest_2018}, there is little empirical evidence of their effectiveness and the drivers of their success. Moreover, TURF-reserve systems are inherently intricate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{ostrom_2009,gelcich_2015}. It is therefore important to consider not only the indicators of interest, but also the governance settings under which the reserves operate.

Recent changes in fisheries regulation in Mexico provide a ripe opportunity to study the effectiveness of community-based TURF-reserves in small-scale fisheries. In Mexico, a legal framework created in 2012 allows fishers to request legal recognition of community-based reserves as ``Fish Refuges'' (\emph{Zona de Refugio Pesquero}, described in more detail below; \citet{nom}). Since 2012, 45 old and new marine reserves have gained legal recognition as Fish Refuges. Of these, 18 were originally implemented within TURFs. However, their effectiveness has not yet been formally evaluated and reported in the scientific literature.

@@

Community-based marine reserves that are implemented within TURFs are a form of TURF-reserve, voluntarily established and enforced by local communities. This bottom-up approach can increase compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the \emph{kapu} or \emph{ra'ui} areas in the Pacific Islands \citep{johannes_2002,bohnsack_2004}. However, community-based reserves can be hard to enforce if they are not legally recognized. In such conditions, TURF fishers must rely on the exclusive access of the TURF to maintain high levels of compliance.

@@

The most common setup of community-based TURF-reserves in Mexico is the following. Fishers from a given community are assembled in fishing cooperatives which have exclusive fishing rights over a spatially delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig \ref{fig:map}A). Each TURF is exclusively fished by one cooperative, and each community usually hosts no more than one cooperative. The profits from each TURF are shared amongst all fishers from the cooperative. Fishing cooperatives interested in implementing marine reserves work with CSOs to implement marine reserves within their TURFs (\emph{i.e.} TURF-reserves). Fishers then ask the government to grant legal recognition to their TURF-reserves as Fish Refuges after conducting a series of studies outlined in the regulation, namely the creation of a baseline of ecological data and historical performance of the fishery \citep{nom}.

@@

We evaluate three community-based no-take TURF-reserve systems implemented in Mexican TURF-managed fisheries, therefore making them TURF-reserves (Fig \ref{fig:map}A). The first one was created by the \emph{Buzos y Pescadores de la Baja California} fishing cooperative, located in Isla Natividad in the Baja California Peninsula (Fig \ref{fig:map}B). The main fishery in the island is the spiny lobster (\emph{Panulirus interruptus}), but other resources like finfish, sea cucumber, sea urchin, snail, and abalone are also an important source of income. In 2006, the community decided to implement two marine reserves within their fishing grounds. The objective of these reserves was ``to protect and recover stocks of commercially important invertebrate species''; mainly lobster and abalone. The reserves obtained legal recognition in 2018 \citep{dof_website_2018}.

@@

The biological data are collected by members from each community and personnel from the Mexican CSO \emph{Comunidad y Biodiversidad} (\href{www.cobi.org.mx}{COBI}). Trained divers record species richness and abundances of fish and invertebrate species along replicate transects (30 \(\times\) 2 m each) at depths 5-20 m in the reserves and control sites, where a minimum of 4 transects per site are performed \citep{suman_2010-ez,fulton_2018,fulton_2019}. Size structures are also collected during fish surveys. All sites were surveyed annually, and at least once before implementation of the reserves. A summary of sampling effort is shown in the supplementary materials (Tables S1-S2).

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings (Kg) and revenues (MXP) for TURF-managed cooperatives with and without marine reserves. In this case our treated unit are the cooperatives (\emph{i.e.} communities) that have implemented a reserve within their TURF, and the controls are nearby communities that have a TURF but did not implement a reserve. These communities have TURFs adjacent to the TURFs of our focal communities. Cooperatives incorporated in this analysis have similar number of members, belong to larger regional-level Cooperative Federations, and are exposed to the same markets and institutional frameworks, making them plausible controls \citep{mccay_2014,mccay_2017,ayer_2018}. Landings and revenues were aggregated at the cooperative-year level, and revenues were adjusted to represent 2014 values by the Consumer Price Index for Mexico \citep{oecd_2017}. A table with summary statistics for this data is provided in the supplementary materials (Table S3, Figure S5).

@@

The coefficient interpretations remains as for Eq. \ref{eqn:reg_bio}, but in this case the \(Treated\) dummy variable indicates if the community has a reserve (\(Treated = 1\)) or not (\(Treated = 0\)). These regression models allow us to establish a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and site-specific dynamics \citep{depalma_2018}. For each indicator in each community, we fit one model (\emph{e.g.} there are three models for lobster density, one for each community) for a total of 12 biological model fits and four socioeconomic model fits. Model coefficients were estimated via ordinary least-squares and used heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R version 3.5.2 and R Studio version 1.1.456 \citep{R_2018}. All data and code needed to reproduce our analyses are available in a GitHub repository at: \url{https://github.com/jcvdav/ReserveEffect}.

We use the SES framework to evaluate each community and create a narrative that provides context for each community. The use of this framework standardizes our analysis and allows us to communicate our results in a common language across fields by using a set of previously defined variables and indicators. Due to the lack of sufficient information to quantitatively operationalize the social-ecological systems framework for these case studies (as in \citet{leslie_2015-na}), we followed a similar approach to \citet{basurto_2013-oq}, who used the SES framework as a classification system of the available information to qualitatively analyze fisheries systems. We based our variable selection primarily on \citet{leslie_2015-na} and \citet{basurto_2013-oq}, who operationalized and analyzed Mexican fishing cooperatives using this framework, and identified the key variables relevant to fishing cooperatives in Mexico. We also incorporate other relevant variables known to influence reserve performance following \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. Table \ref{table:ses} shows the selected variables, along with definitions and values.

@@

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}). For all years before implementation, the effect sizes are close to zero, indicating that the control and treatment sites have similar pre-treatment trends, suggesting that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Interestingly, the negative effect observed for Isla Natividad on year 5 corresponds to the 2011 hypoxia events \citep{micheli_2012-EU}. The only positive change observed in lobster landings is for Isla Natividad in 2014 (\(p < 0.1\)). The year of post-implementation data for Maria Elena does not show a significant effect of the reserve. Isla Natividad shows higher revenues after the implementation of the reserve, as compared to the control communities. However, these changes are only significant for the third year (\(p < 0.05\)). Full tables with model coefficients are presented in the supplementary materials (Tables S4-S5).

@@

Our qualitative implementation of the social-ecological systems framework allowed us to systematically identify important differences between the case studies' governance systems and incorporate other characteristics of these fisheries neglected during the process of data collection (Table \ref{table:ses}). We find that the analyzed communities share similarities known to foster sustainable resource management and increase reserve effectiveness. For example, fishers operate within clearly outlined TURFs (RS2, GS6.1.4.3) that provide exclusive access to resources and reserves. Along with their relatively small groups (A1 - Number of relevant actors), Isolation (A3), Operational rules (GS6.2), Social monitoring (GS9.1), and Graduated sanctions (GS10.1), these fisheries have solid governance structures that enable them to monitor their resources and enforce rules to ensure sustainable management. In general, success of conservation initiatives depends on the incentives of local communities to maintain a healthy status of the resources upon which they depend \citep{jupiter_2017}. Due to the clarity of access rights and isolation, the benefits of conservation directly benefit the members of the fishing cooperatives, which have favored the development of efficient community-based enforcement systems. However, our SES analysis also highlights factors that might hinder reserve performance or mask outcomes. While total reserve size ranges from 0.2\% to 3.7\% of the TURF area, individual reserves are often small (RS3); the largest reserve is only 4.37 km \textsuperscript{2}, and the smallest one is 0.09 km \textsuperscript{2}. Reserves are also relatively young (RS5). Additionally, fishers harvest healthy stocks (RS4.1), and it is unlikely that marine reserves will result in increased catches.

@@

Our analyses of socioeconomic indicators has three limitations. First, we only look at landings and revenues by landings for communities with and without TURF-reserves. There are a number of other possible indicators that could show a change due to the implementation of the reserve. Notably, one often cited in the literature is additional benefits, such as tourism \citep{viana_2017}. However, it is unlikely that the evaluated communities will experience tourism benefits due to their remoteness and the lack of proper infrastructure to sustain tourism. A second limitation of our socioeconomic analysis is that we do not observe effort data, which may mask the effect of the reserve. For example, if catches remain relatively unchanged but fishing effort decreased, that would imply a larger catch per unit effort and thus higher profitability, provided that cost per unit effort does not increase. Likewise, it is possible that fishing effort increased around reserves to maintain the historical levels of landings. A final limitation applies to Maria Elena, where we only observe landings and income for one year after reserve implementation. While one would not expect to observe increased landings or income in such a short period, a spatial closure might cause total catches to decline, especially if effort is held constant.

A first possible explanation for the lack of effectiveness may be the young age of the reserves. Literature shows that age and enforcement are important factors that influence reserve effectiveness \citep{edgar_2014-UO,babcock_2010}. Isla Natividad has the oldest reserves, and our SES analysis suggests that all communities have a well-established community-based enforcement system. With these characteristics, one would expect the reserves to be effective. Maria Elena and Punta Herrero are relatively young reserves (\emph{i.e.} \textless{} 6 years old; RS5 in Table \ref{table:ses}) and effects may not yet be evident due to the short duration of protection, relative to the life histories of the protected species; community-based marine reserves in tropical ecosystems may take six years or more to show a spillover effect \citep{dasilva_2015-zX}.

Another key condition for effectiveness is reserve size \citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be attributed to poor ecological coherence in reserve design (\emph{sensu} \citet{rees_2018}). Previous research has shown that reserves in Isla Natividad yield fishery benefits for the abalone fishery \citep{rossetto_2015-V0}, however, abalone are less mobile than lobsters, and perhaps the reserves provide enough protection to these sedentary invertebrates, but not lobsters. Design principles developed by \citet{green_2017} for marine reserves in the Caribbean state that reserves ``should be more than twice the size of the home range of adults and juveniles'', and suggest that reserves seeking to protect spiny lobsters should have at least 14 km across. As shown through the SES analysis, the size of the marine reserves appears small compared to the movement capacity of the main targeted species (RU1, RS3; Table \ref{table:ses}). Furthermore, fishers may favor implementation of reserves that pose low fishing costs due to their small size or location. Our analysis of economic data supports this hypothesis, as neither landings nor revenues showed the expected short-term reductions associated to the first years of reserve implementation \citep{ovando_2016-Wg}.

@@

\caption{\label{tab:unnamed-chunk-10}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}{l|l}
\hline
Indicator & Units\\
\hline
\multicolumn{2}{l}{\textbf{Biological}}\\
\hline
\hspace{1em}Lobster density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Invertebrate density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish density & org $\mathrm{m}^{-2}$\\
\hline
\hspace{1em}Fish biomass & Kg $\mathrm{m}^{-2}$\\
\hline
\multicolumn{2}{l}{\textbf{Socioeconomic}}\\
\hline
\hspace{1em}Income from target species & M MXP\\
\hline
\hspace{1em}Landings from target species & Metric Tonnes\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{tab:unnamed-chunk-11}\label{table:ses}Variables for the Social-Ecological System analysis (IN = Isla Natividad, ME = Maria Elena, PH = Punta Herrero). Alphanumeric codes follow \citet{basurto_2013-oq}; an asterisk (*) denotes variables incorporated based on \citet{difranco_2016-Xw} and \citet{edgar_2014-UO}. The presented narrative applies equally for all communities unless otherwise noted.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{>{\raggedright\arraybackslash}p{6.5cm}|>{\raggedright\arraybackslash}p{12.2cm}}
\hline
Variable & Narrative\\
\hline
\multicolumn{2}{l}{\textbf{Resource System (RS)}}\\
\hline
\hspace{1em}RS2 - Clarity of system boundaries: Clarity of geographical boundaries of TURF and reserves & Individual TURF and reserve boundaries are explicitly outlined in official documents that include maps and coordinates. Reserve placement is decided by the community. Fishers use GPS units and landmarks.\\
\hline
\hspace{1em}RS3 - Size of resource system: TURF Area (Km$^2$) & IN = 889.5; ME = 353.1; PH = 299.7\\
\hline
\hspace{1em}RS3 - Size of resource system: Reserve area (Evaluated reserve area; Km$^2$) & IN = 2 (1.3); ME = 10.48(0.09); PH = 11.25 (4.37)\\
\hline
\hspace{1em}RS4.1 - Stock status: Status of the main fishery & Lobster stocks are well managed, and are (IN) or have been (ME, PH) MSC certified.\\
\hline
\hspace{1em}*RS5 - Age of reserves: Years since reserves were implemented & IN = 12; ME = 6; PH = 5\\
\hline
\multicolumn{2}{l}{\textbf{Resource Unit (RU)}}\\
\hline
\hspace{1em}RU1 - Resource unit mobility & Adult spiny lobsters can move between 1 and 10 Km, while larvae can have displacements in the order of hundreds of Km (Aceves-Bueno et al., 2017; Green et al., 2017).\\
\hline
\hspace{1em}RU5 - Number of units (catch diversity): Number of targeted species & Lobster is their main fishery of these three communities, but they also target finfish (2 spp each). Additionally, fishers from Isla Natividad target other sedentary benthic invertebrates (4 spp).\\
\hline
\multicolumn{2}{l}{\textbf{Actors (A)}}\\
\hline
\hspace{1em}A1 - Number of relevant actors: Number of fishers & IN = 98; ME = 80; PH = 21\\
\hline
\hspace{1em}*A3 - Isolation: Level of isolation of the fishing grounds & Their fishing grounds and reserves are highly isolated and away from dense urban centers. IN lies 545 Km south from Tijuana, and ME and PH 230 Km south from Cancun, where the nearest international airports are located.\\
\hline
\multicolumn{2}{l}{\textbf{Governance system (G)}}\\
\hline
\hspace{1em}GS6.1.4.3 - Territorial use communal rights : Presence of institutions that grant exclusive harvesting rights & Each community has exclusive access to harvest benthic resources, including lobster. These take the form of Territorial User Rights for Fisheries granted by the government to fishing cooperatives.\\
\hline
\hspace{1em}GS6.2 - Operational rules: Rules implemented by individuals atuhorized to partake on collective activities & Fishers have rules in addition to what the legislation mandates. These are: larger minimum catch sizes, lower quotas, and assigning fishers to specific fishing grounds within their TURF.\\
\hline
\hspace{1em}GS9.1 - Social monitoring: Monitoring of the activities performed by cooperative members and external fishers & Fishing cooperatives have a group (Consejo de vigilancia) that monitors and enforces formal and internal rules. They ensure fishers of their fishing cooperative adhere to the established rules, and that foreign vessels do not poach their TURF and reserves.\\
\hline
\hspace{1em}GS9.2 - Biophysical monitoring: Monitoring of biological resources, including targeted species & Fishers perform annual standardized underwater surveys in the reserves and fishing grounds. Recently, they have installed oceanographic sensors to monitor oceanographic variables.\\
\hline
GS10.1 - Graduated sanctions & Fishers have penalties for breaking collective-choice rules or fishing inside the reserves. These may range from scoldings and warnings to not being allowed to harvest a particular resource or being expelled from the cooperative.\\
\hline
\end{tabular}}
\end{table}
